Carbonate chemistry at Ny-Ålesund

### 2021-03-19 awipev-CO2.Rmd
Finalized choice of K1K2

#################### Regression function
# function regression plot with equation (MA or OLS) in title
ggreg <- function (xdata, ydata, type = "OLS", plot = TRUE, title = "title", subtitle = "subtitle", xlab = "x", ylab = "y") { # x and y are the names of the variables
  fit_data <- tibble(xdata, ydata)
  colnames(fit_data) = c("xdata", "ydata")
  fit <- print(lmodel2::lmodel2(ydata ~ xdata))
if (type == "OLS") ty = 1 #one selects OLS only
if (type == "MA") ty = 2 #one selects MA only
#reg <- fit$regression.results[ty,] #one selects OLS only
intercept <- fit$regression.results$Intercept[ty]
slope <- fit$regression.results$Slope[ty]
if (plot == TRUE) {
  plt <- ggplot(data = fit_data, aes(x = xdata, y = ydata)) + 
  geom_point(na.rm = TRUE, size = 2, colour = "blue") +
  geom_abline(aes(intercept = intercept, slope = slope),
              colour = "blue")  + 
  labs(title = paste(title, "; type: ", type,
                     "\n R2 = ", round(fit$rsquare, digits = 2),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "; P =", signif(fit$P.param, 3)),
       subtitle = subtitle,
       x = xlab,
       y = ylab)
  return(plt)
  } else {
  return(fit)
}
}

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(
                     closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     sal_mix,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     pressure_insitu_ctd
                   )
                 ), by = "closest_datetime") %>% 
  dplyr::select(-closest_datetime)

@@

## Salinities
Several salinity sensors are on site, 1 inside the Ferrybox (SBE45) and 5 outside (ctd183, ctd181, ctd578, ctd964/964b and ctd103) that are switched periodically. Salinity from FB is taken in consideration. If data gaps are encountered, they are filled with bottom salinty data from ctd (> 8 m). New column "sal_mix" is created and is the one used in ESSD paper.

@@

## Impact of choice of constants  {.tabset .tabset-pills}
### Lueker vs Sulpis using means of monthly means
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
The means used are:
`r knitr::kable(tibble(mean_at, mean_ct, mean_pco2, mean_sal, mean_temp), digits = c(1, 1, 0, 1, 1), caption = "Mean discrete values.", format = "simple")`

@@

### Lueker vs Sulpis using FerryBox pCO2 and at_calc
```{r impact of constants on FerryBox data, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at_calc
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at_calc, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_dur_temp_4, pco2_fb, at_calc, sal_fb, temp_fb)
# Using Lueker  
carb_l <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "l",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize Lueker-derived pH at 4°C
carb_l$pH <- pHinsi(pH = carb_l$pH, 
                         ALK = carb_l$ALK, Tinsi = 4, 
                         Tlab = carb_l$T, Pinsi = 0, 
                         S = carb_l$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# Using Sulpis
carb_s <-
  carb(
    24,
    tmp$pco2_fb,
    tmp$at_calc * 1e-6,
    S = tmp$sal_fb,
    T = tmp$temp_fb,
    P = 0,
    Pt = 0,
    Sit = 0,
    k1k2 = "s20",
    kf = "dg",
    ks = "d",
    pHscale = "T",
    b = "u74"
  )
# normalize Sulpis-derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, 
                         ALK = carb_s$ALK, Tinsi = 4, 
                         Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# calculate difference Lueker - Sulpis
diff_ls <- as_tibble(carb_l - carb_s) %>% 
  dplyr::rename(delta_ph = pH, delta_ct = DIC, delta_OmegaAragonite = OmegaAragonite) %>%
  dplyr::select(delta_ph, delta_ct, delta_OmegaAragonite)
# calculate descriptive statistics
diff_ls_quant <- tibble(pH = as.numeric(summary(diff_ls$delta_ph)),
                        CT = as.numeric(summary(diff_ls$delta_ct * 1e6)),
                        OmegaAragonite = as.numeric(summary(diff_ls$delta_OmegaAragonite))) %>%
  dplyr::mutate(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), .before = 1)
# prepare tibble to compare calculated pH and spec pH
tmp2 <- tibble(ph_s_dur_temp_4 = tmp$ph_s_dur_temp_4, pH_l = carb_l$pH, pH_s = carb_s$pH) %>% 
  tidyr::drop_na(ph_s_dur_temp_4)
# summary of differences
diff_lueker <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_l))
diff_sulpis <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_s))
lueker_sulpis_table <- tibble(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), diff_lueker, diff_sulpis)
# make plots
p1 <- ggreg(xdata = tmp2$ph_s_dur_temp_4, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 
            title = "Lueker vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_dur_temp_4, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 
            title = "Sulpis vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis") 
p <- p1 + p2
ggsave(filename = "figures/lueker-sulpis.png", plot = p, width = 20, height = 15, units = "cm")
```
The discrete values of pCO2, AT (calculated from the AT-salinity relationship), salinity and temperature in the FerryBox were used to calculate pH from measured pCO2 and AT (from a salinity-AT relationship) using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). The differences induced by the choice of the carbonic acid constants is:
+ large for pH (0.026 pH units on average; always higher with Sulpis than with Lueker)
+ large for CT (3.7 umol/kg on average; always lower with Sulpis than with Lueker)
+ negligible for OmegaAragonite
`r knitr::kable(diff_ls_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Sulpis using discrete Ferrybox data.", format = "simple")`

Here we compare pH measured spectrophotometrically with pH calculated from pCO2 and AT using the sets of K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). All pH values were normalized to a temperature of 4°C. The absloute differences are large, up to 0.14 pH units. On average, both formulations overestimate spectrophotometric pH, by 0.016 pH units with Lueker et al. (2020) and by 0.042 pH units with Sulpis et al. (2020).
`r knitr::kable(lueker_sulpis_table, digits = 3, caption = "Difference between spectrophotometric pH and pH calculated with pCO2 and AT using the formulations for K1 and K2 of Lueker and Sulpis.", format = "simple")`

>>>>

#################### Regression function
# function regression plot with equation (MA or OLS) in title
ggreg <- function (xdata, ydata, type = "OLS", plot = TRUE, title = "title", subtitle = "subtitle", xlab = "x", ylab = "y") { # x and y are the names of the variables
  fit_data <- tibble(xdata, ydata)
  colnames(fit_data) = c("xdata", "ydata")
  fit <- print(lmodel2::lmodel2(ydata ~ xdata))
if (type == "OLS") ty = 1 #one selects OLS only
if (type == "MA") ty = 2 #one selects MA only
#reg <- fit$regression.results[ty,] #one selects OLS only
intercept <- fit$regression.results$Intercept[ty]
slope <- fit$regression.results$Slope[ty]
if (plot == TRUE) {
  plt <- ggplot(data = fit_data, aes(x = xdata, y = ydata)) + 
  geom_point(na.rm = TRUE, size = 2, colour = "blue") +
  geom_abline(aes(intercept = intercept, slope = slope),
              colour = "blue")  + 
  labs(title = paste(title, "; type: ", type,
                     "\n R2 = ", round(fit$rsquare, digits = 2),
                     "; Intercept =", signif(intercept, 3),
                     "; Slope =", signif(slope, 3),
                     "; P =", signif(fit$P.param, 3)),
       subtitle = subtitle,
       x = xlab,
       y = ylab)
  return(plt)
  } #else {
  #return(fit)
#}
}

@@

# add sal_fb + sal_insitu_ctd + temp_fb + temp_insitu_ctd from d_all into d, to use pHinsi
# use the closest time to add them
closest <- which.closest(d$datetime, d_all$datetime) # closest date and time
d$closest_datetime <- d_all$datetime[closest]
# if the closest value is more than 30 min away, then we do not take it
d$closest_datetime[abs(d$datetime - d_all$datetime[closest]) > 30*60] <- NA
# create closest_datetime in d_all (copy of datetime)
d_all$closest_datetime <- d_all$datetime
d <- left_join(d, d_all %>%
                 dplyr::select(
                   c(closest_datetime,
                     sal_fb,
                     temp_fb,
                     temp_mix,
                     sal_mix,
                     temp_insitu_11m,
                     temp_insitu_ctd,
                     pressure_insitu_ctd
                   )
                 ), by = "closest_datetime") %>% 
  dplyr::select(-closest_datetime)

@@

## Salinities
Several salinity sensors are on site, 1 inside the Ferrybox (SBE45) and 5 outside (ctd183, ctd181, ctd578, ctd964/964b and ctd103) that are switched periodically. Salinity from FB is taken in consideration. If data gaps are encountered, they are filled with bottom salinity data from ctd (> 8 m). New column "sal_mix" is created and is the one used in ESSD paper.

@@

## Impact of choice of constants  {.tabset .tabset-pills}
### Lueker vs Sulpis using mean discrete values
```{r impact of constants discrete, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using mean discrete values
# first, order columns
d_all <- d_all %>% 
  dplyr::select(order(colnames(d_all))) %>% 
  dplyr::select(datetime, date, year, monthd, montht, everything())
# calculate means of discrete values
mean_ct <- mean(discrete$ct, na.rm = TRUE)
mean_at <- mean(discrete$at, na.rm = TRUE)
mean_pco2 <- mean(discrete$pco2_calc_fb, na.rm = TRUE)
mean_sal <- mean(discrete$sal_fb, na.rm = TRUE)
mean_temp <- mean(discrete$temp_fb, na.rm = TRUE)
mean_discrete <- tibble(mean_ct = mean_ct, mean_at = mean_at, mean_pco2 = mean_pco2, mean_sal = mean_sal, mean_temp = mean_temp)
# calculate typical uncertainty assuming uncertainties of 3 uatm for pCO2 and 4 umol/kg for AT and CT
errors_pco2_at <- seacarb::errors(flag=24, var1=mean_pco2, var2=1e-6*mean_at, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=3, evar2 = 4e-6)
errors_at_ct <- seacarb::errors(flag=15, var1=1e-6*mean_at, var2=1e-6*mean_ct, S=mean_sal,
                          T=mean_temp, P=0, Pt=0, Sit=0, k1k2="l", kf="dg", ks="d", pHscale="T",
                          b="u74", evar1=4e-6, evar2 = 4e-6)
errors <- tibble(e_pH = c(errors_pco2_at$pH, errors_at_ct$pH),
                 e_pCO2 = c(NA, errors_at_ct$pCO2),
                 e_DIC = c(errors_pco2_at$pH, NA)*1e3,
                 e_Oa = c(errors_pco2_at$OmegaAragonite, errors_at_ct$OmegaAragonite)
)
# impacts using pco2 and AT pair
carb_l_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_1 <- carb(24, mean_pco2, mean_at*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_1 <- carb_l_1 - carb_s_1
# impacts using AT-CT pair
carb_l_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0, 
                 Pt = 0, Sit = 0,
                 k1k2 = "l", kf = "dg", ks = "d", pHscale = "T", b = "u74")
carb_s_2 <- carb(15, mean_at*1e-6, mean_ct*1e-6, S = mean_sal, T = mean_temp, P = 0,
                Pt = 0, Sit = 0, 
                k1k2 = "s20", kf = "dg", ks = "d", pHscale = "T", b = "u74")
diff_ls_2 <- carb_l_2 - carb_s_2
# summary
const_choice <- bind_rows(diff_ls_1, diff_ls_2) %>% 
  dplyr::mutate(flag = c("pCO2 and ALK", "ALK and DIC"),
                S = mean_sal,
                T = mean_temp,
                DIC = DIC * 1e6
                ) %>% 
  dplyr::select(flag, pH, pCO2, DIC, OmegaAragonite) %>% 
  dplyr::bind_cols(errors)
```
The means used are:
`r knitr::kable(mean_discrete, digits = c(1, 1, 0, 1, 1), caption = "Mean discrete values.", format = "simple")`

@@

### Lueker vs Sulpis using FerryBox pCO2 and at_calc
```{r impact of constants on FerryBox data, echo=FALSE, warning=FALSE, message=FALSE}
# Compare Lueker and Sulpis using pco2_fb and at in situ
tmp <- d_all %>% 
  tidyr::drop_na(pco2_fb, at, sal_fb, temp_fb) %>% 
  dplyr::select(ph_s_dur_temp_4, pco2_fb, at_calc, sal_fb, temp_fb)
# Using Lueker  
carb_l <- carb(flag = 24, var1 = tmp$pco2_fb, var2 = tmp$at_calc * 1e-6, S = tmp$sal_fb, 
               T = tmp$temp_fb, P = 0, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks = "d", 
               pHscale = "T", b = "u74")
# normalize Lueker-derived pH at 4°C
carb_l$pH <- pHinsi(pH = carb_l$pH, ALK = carb_l$ALK, Tinsi = 4, Tlab = carb_l$T, Pinsi = 0, 
                         S = carb_l$S, Pt = 0, Sit = 0, k1k2 = "l", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# Using Sulpis
carb_s <- carb(flag = 24, var1 = tmp$pco2_fb, var2 = tmp$at_calc * 1e-6, S = tmp$sal_fb, 
               T = tmp$temp_fb, P = 0, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks = "d", 
               pHscale = "T", b = "u74")
# normalize Sulpis-derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, ALK = carb_s$ALK, Tinsi = 4, Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "s20", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# Using Papadimitriou
carb_p <- carb(flag = 24, var1 = tmp$pco2_fb, var2 = tmp$at_calc * 1e-6, S = tmp$sal_fb, 
               T = tmp$temp_fb, P = 0, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks = "d", 
               pHscale = "T", b = "u74")
# normalize Papadimitriou-derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, ALK = carb_s$ALK, Tinsi = 4, Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "p18", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# Using Millero et al. (2002)
carb_m <- carb(flag = 24, var1 = tmp$pco2_fb, var2 = tmp$at_calc * 1e-6, S = tmp$sal_fb, 
               T = tmp$temp_fb, P = 0, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks = "d", 
               pHscale = "T", b = "u74")
# normalize Millero et al. (2002)-derived pH at 4°C
carb_s$pH <- pHinsi(pH = carb_s$pH, ALK = carb_s$ALK, Tinsi = 4, Tlab = carb_s$T, Pinsi = 0, 
                         S = carb_s$S, Pt = 0, Sit = 0, k1k2 = "m02", kf = "dg", ks="d", 
                         pHscale = "T", b="u74")
# calculate difference Lueker - Sulpis
diff_ls <- as_tibble(carb_l - carb_s) %>% 
  dplyr::rename(delta_ph = pH, delta_ct = DIC, delta_OmegaAragonite = OmegaAragonite) %>%
  dplyr::select(delta_ph, delta_ct, delta_OmegaAragonite)
# calculate descriptive statistics Lueker - Sulpis
diff_ls_quant <- tibble(pH = as.numeric(summary(diff_ls$delta_ph)),
                        CT = as.numeric(summary(diff_ls$delta_ct * 1e6)),
                        OmegaAragonite = as.numeric(summary(diff_ls$delta_OmegaAragonite))) %>%
  dplyr::mutate(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), .before = 1)
# prepare tibble to compare calculated pH and spec pH
tmp2 <- tibble(ph_s_dur_temp_4 = tmp$ph_s_dur_temp_4, pH_l = carb_l$pH, pH_s = carb_s$pH, pH_p = carb_p$pH, pH_m = carb_m$pH) %>% 
  tidyr::drop_na(ph_s_dur_temp_4)
# summary of differences
lueker <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_l))
sulpis <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_s))
papadimitriou <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_p))
millero2002 <- as.numeric(summary(tmp2$ph_s_dur_temp_4 - tmp2$pH_m))
diff_table <- tibble(var = c("Min", "Q1", "Median", "Mean", "Q3", "Max"), lueker, sulpis, papadimitriou, millero2002)
# make plots
p1 <- ggreg(xdata = tmp2$ph_s_dur_temp_4, ydata = tmp2$pH_l, type = "MA", plot = TRUE, 
            title = "Lueker vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Lueker")
p2 <- ggreg(xdata = tmp2$ph_s_dur_temp_4, ydata = tmp2$pH_s, type = "MA", plot = TRUE, 
            title = "Sulpis vs measured", subtitle = "both pH normalized at 4°C",
            xlab = "Spectrophotometric pH measured @ 4°C",
            ylab = "pH @ 4°C calculated (pCO2+AT) with Sulpis") 
p <- p1 + p2
ggsave(filename = "figures/lueker-sulpis.png", plot = p, width = 20, height = 15, units = "cm")
The discrete values of pCO2, salinity and temperature in the FerryBox and in situ AT were used to calculate pH from measured pCO2 and AT using the formulations for K1 and K2 of Lueker et al. (2000) and Sulpis et al. (2020). The differences induced by the choice of the carbonic acid constants is:

- small for pH (`r round(diff_ls_quant[4, 2], digits = 3)` pH units on average; ; median is `r round(diff_ls_quant[3, 2], digits = 3)`) compared with the uncertainty calculated with seacarb (`r round(const_choice[1, 6], digits = 3)`. In general, pH is higher with Sulpis than with Lueker.
- very good for CT (`r round(diff_ls_quant[4, 3], digits = 1)` umol/kg on average; always lower with Sulpis than with Lueker) compared with the uncertainty calculated with seacarb (`r round(const_choice[1, 8], digits = 1)`)
- negligible for OmegaAragonite

`r knitr::kable(diff_ls_quant, digits = c(0, rep(3, 5)), caption = "Lueker minus Sulpis using discrete Ferrybox data.", format = "simple")`

Here we compare pH measured spectrophotometrically with pH calculated from pCO2 and AT using various formulations of K1 and K2. All pH values were normalized to a temperature of 4°C. The absolute differences are large, up to `r round(max(abs(diff_table[, 2:5])), digits=3)` pH units. On average, both formulations overestimate spectrophotometric pH. The calculated pH closer to measured pH is using the formulation of Lueker et al. (2020) with a mean difference of `r round(max(diff_table[4, 2:5]), digits=3)` pH units. This difference is twice larger than the uncertainty for pH calculated from pCO2 and AT according to seacarb (`r round(errors_pco2_at$pH, digits=3)`).
`r knitr::kable(diff_table, digits = 3, caption = "Difference between spectrophotometric pH and pH calculated with pCO2 and AT using different formulations for K1 and K2.", format = "simple")`
`r include_graphics("figures/lueker-sulpis.png")`
